---
output:
  pdf_document: 
    fig_width: 7
    fig_height: 7
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r include=FALSE}
library(readr)
library(prora)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE, include=FALSE)
```

```{r}
genomics <- readr::read_tsv("Count_QC-rpkm_V2.txt")
head(genomics)
proteomics <- readxl::read_xlsx("second_Draft_V4_WithContaminants/path_qc/transformed_ProteinIntensities_WIDE.xlsx")
```

Some protein_Id have more than 2 entries - we separate them into rows generating duplicated measurements.

```{r}
proteomics$row_Id <- 1:nrow(proteomics)
dim(proteomics)
prot2 <- proteomics %>% tidyr::separate_rows(protein_Id, sep = ";")
dim( prot2 )
```

Here, are the duplicated rows:

```{r}
xx <- table(prot2$row_Id)
xx <- as.array(xx)
xx <- xx[xx > 1]
duplicated <- as.numeric(names(xx))
(proteomics[duplicated[1],])$protein_Id

```


#quantable::write.vector(proteomics$protein_Id, file = "tmp.txt")

Now we map the uniprot Id's to ensemble. Some of the uniprot ID's have multiple ensemble id's assigned (up to 14)


```{r}
prot_mapped_table <- prora::map_ids_uniprot(prot2,ID_col = "protein_Id", to = "ENSEMBL_ID")
plot(table(prot_mapped_table$protein_Id))
xx <- sort(as.array(table(prot_mapped_table$protein_Id)), decreasing = TRUE)
(xx[xx > 1])

```

Some of the uniprot ID's were not mapped to ensemble ID's.

```{r}
prot_mapped_table$protein_Id[is.na(prot_mapped_table$ENSEMBL_ID)]

```


Now we check which of the ensemble ids, obtained using the uniprot mapping service, are also present in the genomic data.

The table below lists all uniprot ID's which are not mapped to genomic data. NA-indicates that no Ensemble ID mapping was available for the uniprot ID, while the shown ensemble ID's are not present in the genomic data.
I am a bit surprised why so many Ensemble_ID's could not be found in the genomic data although we have a corresponding protein product (441 Ensemble Id's). Is there a way to check if those Ensemble  ID's are recently introduced, 
while the db you used is older?

```{r}
sum(!na.omit(prot_mapped_table$ENSEMBL_ID) %in% genomics$`Feature ID`)
```

```{r eval=FALSE, include=FALSE}
knitr::kable(prot_mapped_table[!prot_mapped_table$ENSEMBL_ID %in% genomics$`Feature ID`, ] %>% select("protein_Id", "ENSEMBL_ID"))

```


```{r}
cnamprot <- gsub("CMP/MEP","CMP_MEP",colnames(prot_mapped_table)[4:ncol(prot_mapped_table)])
cnamprot <- gsub("Prt","PID",cnamprot)
cnamprot <- gsub("757/767PID","PID757/767",cnamprot)
cnamprot <- gsub("304PID","PID304", cnamprot)
cnamprot <- gsub("240PID", "PID240", cnamprot)

grep("CMP_MEP~control.NO.no~Haemo", cnamprot)

cnamprot <- gsub("CMP_MEP~control.NO.no~Haemo","CMP_MEP~control.NO.no~PID392/732/777/790", cnamprot)
cnamprot <- gsub("CMP_MEP~patient.NO.no~PID64","CMP_MEP~patient.NO.no.~PID67/206/293", cnamprot)
cnamprot <- gsub("HSC~control.NO.no~Haemo", "HSC~control.NO.no~PID392/732/777/790", cnamprot)
cnamprot <- gsub("HSC~patient.NO.no~PID64", "HSC~patient.NO.no~PID67/206/293", cnamprot)

mean(duplicated(cnamprot))

colnames(prot_mapped_table)[4:ncol(prot_mapped_table)] <- cnamprot

cnamgen <- colnames(genomics)[-1]
cnamprot[!cnamprot %in% cnamgen]

cnamgen[!cnamgen %in% cnamprot]



prot_mapped_table %>% select("protein_Id", "ENSEMBL_ID") %>% duplicated %>% mean

prot_mapped_table <- prot_mapped_table %>% select(-one_of("PG.ProteinAccessions","row_Id"))


prot_long <- tidyr::gather(prot_mapped_table, key = "sample",
  value = "medpolish", -(1:2) )

gene_long <- tidyr::gather(genomics, key = "sample",
  value = "intensity", -1 )

colnames(gene_long)[1] <- "ENSEMBL_ID"

gene_long %>% select(ENSEMBL_ID, sample) %>% duplicated %>% mean()
prot_long %>% select(ENSEMBL_ID, sample) %>% duplicated %>% mean()
#gene_long$ENSEMBL_ID <- as.numeric(sub("^ENSG0+","", gene_long$ENSEMBL_ID))
both_data <- inner_join(prot_long, gene_long)
dim(both_data)
both_data %>% select(ENSEMBL_ID, sample) %>% duplicated %>% mean()
both_data %>% select(ENSEMBL_ID, protein_Id, sample) %>% duplicated %>% mean()

both_data <- both_data %>% unite("E_P", ENSEMBL_ID, protein_Id , sep = "~")
geneWide  <- both_data %>% select(-one_of("medpolish")) %>% tidyr::spread(sample , intensity )
geneMat <- as.matrix(geneWide[,-1])
rownames(geneMat) <- geneWide[[1]]

protWide <- both_data %>% select(-one_of("intensity")) %>%
  tidyr::spread(sample , medpolish )
protMat <- as.matrix(protWide[,-1])
rownames(protMat) <- protWide[[1]]
```

# Protein mRNA correlations

```{r proteinCors, fig.cap="density of protein mRNA correlation", include=TRUE}
corRow <- function(i, protMat, geneMat){
  res <- cor(protMat[i,], geneMat[i,], use = "pairwise.complete.obs",
             method = "spearman")
  return(res)
}

protCors <- sapply(1:nrow(protMat), corRow, protMat, geneMat)
length(protCors)
names(protCors) <- rownames(protMat)

plot(density(na.omit(protCors)))
abline(v = 0, col = 2 )
abline(v = summary(na.omit(protCors)), col = "grey")
summary(na.omit(protCors))
```

```{r saveCorrelations}
protCorsTab <- as_tibble(protCors, rownames = "row.names")
writexl::write_xlsx(protCorsTab,"protCorsTab.xlsx")

```


```{r fig.cap="negative correlation", include=TRUE}
anti <- names(sort(protCors)[1:10])

par(mfrow = c(2,2))
plot(protMat[anti[1],], geneMat[anti[1],], main = anti[1])
plot(protMat[anti[2],], geneMat[anti[2],], main = anti[2])
plot(protMat[anti[3],], geneMat[anti[3],], main = anti[3])
plot(protMat[anti[4],], geneMat[anti[4],], main = anti[4])

```


```{r fig.cap="positive correlation", include=TRUE}
anti <- names(sort(protCors, decreasing = TRUE)[1:10])
par(mfrow = c(2,2))
plot(protMat[anti[1],], geneMat[anti[1],], main = anti[1])
plot(protMat[anti[2],], geneMat[anti[2],], main = anti[2])
plot(protMat[anti[3],], geneMat[anti[3],], main = anti[3])
plot(protMat[anti[4],], geneMat[anti[4],], main = anti[4])

```



# Sample Correlations

```{r }

res <- cor(protMat, geneMat, use = "pairwise.complete.obs", method = "spearman")


```


```{r hmmap1,fig.caption= "sample correlation heatmap", fig.width=10, fig.height=10, include=TRUE}
(pheatmap::pheatmap(res, cluster_cols = FALSE,
                         cluster_rows = FALSE,
                         color = colorRampPalette(rev(RColorBrewer::brewer.pal(n = 7, name =
  "YlGnBu")))(100)))

```


```{r hmmap2, fig.caption= "sample correlation heatmap with clustering", fig.width=10, fig.height=10, include=TRUE}
(pheatmap::pheatmap(res ,
                    color = colorRampPalette(rev(RColorBrewer::brewer.pal(n = 7, name =
  "YlGnBu")))(100)))


```
